---
title: "Introduction to textured"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction-to-textured}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The textured package is meant for semi-struct*ured* *text* data, i.e. data that was contained within free text, but still had some structure.


## A sample scenario

This package was developed to help extract information from patient letters from a hospital, where the letters had some structure, but was inconsistent between one letter to the text.

Here's an example letter:

```
Ref: AA/11/BBBB
Date Dictd: 09/01/2019
Date Typed: 15/01/2019

Dear Dr. A Andrews

Mr. Bob BURNQUIST D.O.B. 10/10/1976

Date/Time of Appt: 09/01/2019
Clinic: Rheumatology

Reason for attendance:
Follow-up - rheumatoid arthritis

Rheumatological Diagnoses
Inflammatory spondyloarthritis

Non-rheumatological Diagnoses
Osteoporosis

Medications
Sulfasalazine 105gms bd

Observations in Clinic
Blood pressure 120/71 mmHg today

Assessment
I reviewed...

```

And here's another:

```

Ref: CC/22/GGGG
Date Dictd: 05/05/2019
Date Typed: 12/05/2019

Dear Dr. C Clarkson

Mr. D Dancer D.O.B. 10/10/1955

Date/Time of Appt: 05/05/2019
Clinic: Rheumatology

Reason for attendance:
L1 end plate fracture

Diagnosis
Fibromyalgia
Mild Asthma

Current medications
Omeprazole 20mg twice daily

Observations in Clinic
Blood pressure 120/71mmHg today

Actions for GP / Recommendations
Nil

Management Plan (including Actions for Patients)
1. Bloods checked...

Assessment
I reviewed...
```

While the two letters above are similar, they're slightly different in structure; both have differing sections, and similar sections have different headers. This can make extraction difficult (especially if considering much more letters (e.g. 8,000), as well as storage - a tabular format is not suitable for letters with an inconsistent structure.

However, XML, a semi-structure data structure, is much more suitable, and this is what this package revolves around: the splitting and conversion of documents into an XML format. Once converted, the data can then be read in XML software to be analysed, where there are tools which are much more suitable.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(textured)
```

```{r message=FALSE, warning=FALSE}
sample_1 <- read_csv("data/sample_letters.csv")

sample_1
```

```{r}
headings <- c("Ref:", "Date Dictd:", "Date Typed", "\nDear", "D\\.O\\.B\\.",
              "Date/Time of Appt:", "Clinic:", "Reason for attendance:",
              "Diagnoses:", "Non-rheumatological Diagnoses", "Rheumatological Diagnoses",
              "Current medications:", "Medications:", "Observations in Clinic",
              "Assessment", "Actions for GP / Recommendations",
              "Management Plan \\(including Actions for Patients\\)")
# Note: txx_text_to_txml is sensitive to regular expressions (this is why
# there's escape characters)
```

Performing on a single letter:

```{r}
a <- sample_1$letter_text[1]
b <- txx_text_to_xml(strings = a, tags = headings)
```

```{r}
cat(b)
```

Performing from a dataframe (or tibble):
```{r}
sample_2 <- sample_1 %>%
  mutate(letter_text_xml = txx_text_to_xml(letter_text, tags = headings))

sample_2
```


You can add additional XML elements with `txx_create_element()`.
```{r}
sample_3 <- sample_2 %>%
  mutate(letter_text_xml = paste0(
    txx_create_element(patient_id, "patient"),
    letter_text_xml
  )) %>%
  mutate(letter_xml = txx_create_element(letter_text_xml, "letter"))
```


Then collate the XML elements into a single XML document.
```{r}
letters_xml <- txx_create_element(paste0(sample_3$letter_xml, collapse = ""), "root")
```


You can then process this as an XML document. For example:
```{r message=FALSE, warning=FALSE}
library(XML)
```

```{r message=FALSE, warning=FALSE}
letters_xml <- xmlParse(letters_xml)

letters_df <- xmlToDataFrame(letters_xml, stringsAsFactors = FALSE)

as_tibble(letters_df)
```

```{r}
longer <- letters_df %>%
  pivot_longer(cols = -patient, names_to = "section", values_to = "text")

longer
```





